---
title: "DataViz Homework 04"
author: "Your Name Here"
output: html_document
---

#  Homework 04:  Layering & grouping

***General instructions for homeworks***: 

+  Make a new R Markdown file (.Rmd), referring to the assignment on the course Github page
+  Change the heading to include your author name
+  Save the R Markdown file (named as:  [MikeID]-[Homework01].Rmd -- e.g. "mlopez-Lab01.Rmd") to somewhere where you'll be able to access it later (zip drive, My Documents, Dropbox, etc)
+  Your file should contain the code/commands to answer each question in its own code block, which will also produce plots that will be automatically embedded in the output file
+  **Each answer must be supported by written statements (unless otherwise specified) as well as any code used**: In other words, if the answer is `24`, you should write "The answer is 24" (as opposed to just showing the code and output).
+  Include the names of anyone you collaborated with at the top of the assignment



##  Problem 1

The website Vox.com published an interesting article linked [here](http://www.vox.com/policy-and-politics/2016/10/10/13226264/us-crime-rate-poll). Our interest is in the first graph, titled `Is there more or less crime in the US than two decades ago?`

+ In 2-3 clear and concise sentences, describe what this plot is trying to show
+ Identify the margin of error in the article, and use this metric to consider if the differences in crime type are meaningfull.
+ There are a few ways to improve this chart. Name two.

## Problem 2

Here's a chart of the ratio of airline casualties to passengers over time, linked [here](https://i.redd.it/a8cn7k89xpox.png).

+ In 2-3 clear and concise sentences, describe what this plot is trying to show
+ There are a several ways to improve this chart. Name five.


## Problem 3

Return to the `midwest` data set from `Lab05`.

```{r, warnings = FALSE, message= FALSE}
library(ggplot2)
```

Compare the link between county-level percentage in poverty (y-axis) and education (x-axis), accounting for things like state, population total, and/or race. Use one chart only - there is no right answer!


## Problem 4

Return to the `nycflights` data set. *Note: copy the following code, which drops flights with missing data*.

```{r, warnings = FALSE, message = FALSE}
library(nycflights13)
library(dplyr)
flights <- na.omit(flights)
```

Make the following chart [(link)](https://github.com/statsbylopez/DataViz/blob/master/Homeworks/Homework04_plots.pdf). Which carrier tends to save the most time in the air, based on their average departure and arrival delays?




