---
title: "DataViz Homework 08"
author: "Your Name Here"
output: html_document
---

#  Homework 08:  Making Maps

**Note**: This homework is the combination of a lab and a homework. Answer questions where provided, but feel free to delete any redundant material.

***General instructions for homeworks***: 

+  Make a new R Markdown file (.Rmd), referring to the assignment on the course Github page
+  Change the heading to include your author name
+  Save the R Markdown file (named as:  [MikeID]-[Homework01].Rmd -- e.g. "mlopez-Lab01.Rmd") to somewhere where you'll be able to access it later (zip drive, My Documents, Dropbox, etc)
+  Your file should contain the code/commands to answer each question in its own code block, which will also produce plots that will be automatically embedded in the output file
+  **Each answer must be supported by written statements (unless otherwise specified) as well as any code used**: In other words, if the answer is `24`, you should write "The answer is 24" (as opposed to just showing the code and output).
+  Include the names of anyone you collaborated with at the top of the assignment



## Part I: Additional map notes

On page two of the [maps package documentation](https://cran.r-project.org/web/packages/maps/maps.pdf), we see that in addition to `state`, the `maps` package includes `county`, `france`, `italy`, `nz`, `usa`, `world` and `world2` files.

Let's take a look at the state level data. First, we grab the data and store it as `MainStates`, plotting with a light blue fill. 


```{r eval=FALSE}
library(ggthemes)
library(dplyr)
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
MainStates <- map_data("state")
ggplot() + 
  geom_polygon( data=MainStates, aes(x=long, y=lat, group=group),
                color="black", fill="lightblue", linetype = 1, alpha = 1, size = 1) + 
  coord_fixed(1.3)  + 
  theme_map()
```

There are a few ways to potentially improve this map. 

First, you can change the size of the polygon lines using the `size = `. The default is `size = 1`, and values greater than 1 result in larger line widths. 

Second, `linetype` dictates the type of line with which we make the polygon. The default is straight lines (`alpha = 1`), but these can be tweaked to dashed lines of different types. 

Third, `alpha` dictates the filling density. The default is `alpha = 1` - values closer to 0 will lead to a lighter fill.

Please use this graph above for the following three questions.

**Q1** Use the graph above: Make a plot using a line color of blue, a linetype of 4, a dark-green fill with a filling density of 0.3, and a line width of 0.8. 

**Q2** Use the graph above: Remove the `group = group` command, and explain what happens when the grouping is ignored.

**Q3** Use the graph above: Add in three points, showing (i) Chicago (ii) Saratoga Springs and (iii) one unique place that you have travelled too or would like to travel too. 




## Part II Customizing your map

In addition to data files from the `maps` package, we will use the  [StatePopulation](https://simple.wikipedia.org/wiki/List_of_U.S._states_by_population) data, which includes the **state name**, **estimated population**, and the **number of electoral college votes** each state is allocated.

Now that we have created a base map of the mainland states, we will color each state according to its population. The first step is to use the `dplyr` package to merge the `MainStates` and `StatePopulation` files.

```{r, eval = FALSE}
StatePopulation <- read.csv("https://raw.githubusercontent.com/statsbylopez/DataViz/master/Homeworks/Wiki.States.csv", head = TRUE, sep = ",")
StatePopulation %>% head(3)
StatePopulation1 <- StatePopulation %>% 
  mutate(region = tolower(State))
# Use the dplyr package to merge the MainStates and StatePopulation files
MergedStates <- inner_join(MainStates, StatePopulation1, by = "region")
MergedStates %>% head()
```


**Q4** Identify the role of the `tolower()` function above. Additionally, why did this new variable need to be created?


**Q5** What columns were added to the `MainStates` file when it was joined with the `StatePopulation` file?

Next we create a base map of the mainland United States, using state population as a fill variable.

```{r, eval = FALSE}
# Create a Choropleth map of the United States
p <- ggplot() + 
  geom_polygon(data=MergedStates, 
          aes(x=long, y=lat, group=group, fill = Population/1000000), 
          color="white", size = 0.2) + 
  coord_fixed(1.3)
p
```

**Remarks**

* Instead of using `population`, we use `population/1000000`. Each state is colored by population size (in millions) to make the legend easier to read.

* Once a map is created, it is often helpful to modify color schemes, determine how to address missing values (na.values) and formalize labels. Notice that we assigned the graph a name, `p`. This is particularly useful as we add new components to the graph.

```{r, eval = FALSE}
p1 <- p + scale_fill_continuous(name="Population\n(millions)", 
            low = "lightgreen", high = "darkgreen",limits = c(0,40), 
            breaks=c(5,10,15,20,25,30,35), na.value = "grey50") +
           labs(title="State Population in the Mainland United States") 
p1
```


**Q6** Create a choropleth map showing the number of electorate votes each state gets (`electorate`). Make the state borders purple with size = 1. Also change the color scale for state electorates, with low electorates colored white and states with high populations colored dark red. 



## Part III: the `us.cities` file

The `maps` package also includes a `us.cities` file. The following code adds a point for each major city in the United States. Notice that the size of the point is determined by the population of that city.

```{r, eval = FALSE}
us.cities %>% head()
p <- ggplot() + 
  geom_polygon(data = MainStates, aes(x = long, y = lat, group = group), fill = "grey") + 
        geom_point(data=us.cities, aes(x=long, y=lat, size = pop)) + 
        scale_size(name="Population") + 
  coord_fixed(1.3)
p
```

It appears that the `us.cities` file includes cities in Hawaii and in Alaska. We use the `dplyr` package to eliminate these four cities and make a final set of maps.

```{r, eval = FALSE}
MainCities <- filter(us.cities, long>=-130)

p <- ggplot() + 
  geom_polygon(data=MergedStates, 
          aes(x=long, y=lat, group=group, fill = Population/1000000), 
          color="white", size = 0.2) + 
  coord_fixed(1.3)
  
p1 <- p + scale_fill_continuous(name="State Population", low = "lightblue", 
            high = "darkblue",limits = c(0,40), breaks=c(5,10,15,20,25,30,35), 
            na.value = "grey50") +
      labs(title="Population (in millions) in the Mainland United States")
p1

p2 <- p1 + geom_point(data=MainCities, aes(x=long, y=lat, size = pop/1000000), 
            color = "gold", alpha = .5) + scale_size(name="City Population")
p2
```

**Q7** Restrict the `MainStates` and `MainCities` files to only a few contiguous states, such as New York, Vermont,  New Hampshire, Massachusetts, Rhode Island, and Connecticut. Make a plot of these cities, showing only state capitals. 

You could use the following code for help:

```{r eval=FALSE}
NewStates <- MainStates %>%
  filter(region ==  "new york" | region ==
     "vermont" | region ==  "new hampshire" | region ==
     "massachusetts" | region ==  "rhode island" | 
      region ==  "connecticut")
NewCities <- MainCities %>% 
  filter(capital == 1)
```


